---
title: "pdb2uniprot()"
# author: "Juan Carlos Aledo"
# date: "1/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Returns the uniprot id of a given chain within a PDB structure

### Usage
pdb2uniprot(pdb, chain)

### Arguments
_pdb_ the 4-letter PDB identifier.

_chain_	letter identifying the chain.

### Value
The function returns the UniProt ID for the chain of interest.

### See Also
_uniprot2pdb()_, _id.mapping()_

### Details

The _ptm_ package contains a number of ancillary functions that deal with [Protein Data Bank](https://www.rcsb.org/) (PDB) files. These functions may be useful when structural 3D data need to be analyzed. The mentioned functions are:

* [uniprot2pdb](./uniprot2pdb)
* pdb2uniprot (the current document)
* [pdb.chain](./pdb.chain)
* [pdb.quaternary](./pdb.quaternary)
* [pdb.res](./pdb.res)
* [pdb.pep](./pdb.pep)
* [pdb.select](./pdb.select)
* [download.dssp](./download.dssp)
* [compute.dssp](./compute.dssp)
* [parse.dssp](./parse.dssp)
* [mkdssp](./mkdssp)

One of the most basic operations consists of interconversion between identifiers. For instance, suppose we have a PDB file containing information about the structure of the human deoxyhaemoglobin (2HHB) and we want to know the UniProt IDs of their constitutive polypeptide chains.


```{r}
for (chain in pdb.chain('2hhb')){
  cat(pdb2uniprot('2hhb', chain), '\n')
}
```

We observe that this protein es a heterotetramer formed by two types of polypeptide whose UniProt IDs are given above.

If we don't care about the chains, and all we want is the UniProt IDs for all the polypeptides present in the PDB file, we can use:

```{r}
id.mapping('2hhb', from = 'pdb', to = 'uniprot')
```


