---
title: "acc.dssp()"
# author: "Juan Carlos Aledo"
# date: "1/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
library(knitr)
```

### Description
Computes Residue Accessibility and SASA

### Usage
acc.dssp(pdb, dssp = 'compute', aa = 'all')

### Arguments
_pdb_	is either a PDB id, or the path to a pdb file.

_dssp_	string indicating the preferred method to obtain the dssp file. It must be either 'compute' or 'mkdssp.

_aa_	one letter code for the amino acid of interest, or 'all' for all the protein residues.

### Value
A dataframe where each row is an individual residue of the selected protein. The variables computed, among others, are: (i) the secondary structure (ss) element to which the residue belongs, (ii) the solvent accessible surface area (sasa) of each residue in square angstrom (Ų), and (iii) the accessibility (acc) computed as the percent of the sasa that the residue X would have in the tripeptide GXG with the polypeptide skeleton in an extended conformation and the side chain in the conformation most frequently observed in proteins. For the given PDB, the function obtains its corresponding DSSP file using the chosen method. The argument dssp allows two alternative methods: _compute_ (by default, it calls to the function compute.dssp(), which in turn uses an API to run DSSP at the CMBI) and 
_mkdssp_ (if you have installed DSSP on your system and in the search path for executables).

### References

* Miller et al (1987) [J. Mol. Biol. 196: 641-656](https://www.ncbi.nlm.nih.gov/pubmed/3681970).
* Touw et al (2015) [Nucl. Ac. Res. 43(Database issue): D364-D368](https://www.ncbi.nlm.nih.gov/pubmed/25352545).

### Details

Analysis of protein commonly requires the partition of their structure into regions such as the surface, interior, or interface. Whenever you are interested into analysing such a partition in a particular protein, the following function of the _ptm_ package may be of help:

* acc.dssp (the current document)
* [get.area](/get.area)
* [atom.dpx](/atom.dpx)
* [res.dpx](/res.dpx)
* [stru.part](/stru.part)


The function **acc.dssp()** can be of help to detect changes in accessibilities (ACC) when the analysis shifts its focus from the monomer to the complex. To illustrate this point, let's use the methonine S-adenosyltransferase from _Ureaplasma ureliticum_, which is a homotetramer, as an example. 

```{r}
uMAT <- acc.dssp(pdb = '6rjs')
kable(head(uMAT))
```

If we define the protein-protein interface as the set of residues for which $latex \Delta ACC > 0$, we can provide some basic information about the protein-protein interface. For instance, the fraction of amino acid involved in this protein-protein interface can be computed as:

```{r}
round(length(which(uMAT$delta_acc != 0))/nrow(uMAT), 3) 
```

Or we may know the surface (in Ų) involved in this interface:

```{r}
sum(uMAT$delta_sasa)
```

And so on. 

