---
title: "get.area()"
# author: "Juan Carlos Aledo"
# date: "1/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
library(knitr)
```

### Description
Computes Online Surface Energies Using the Getarea Server.

### Usage
get.area(pdb, keepfiles = FALSE)

### Arguments
_pdb_	is either a PDB id, or the path to a pdb file.

_keepfiles_	logical, if TRUE the dataframe will be saved in the working directory and we will keep the getarea txt file.

### Value
This function returns a dataframe containing the requested information.

### References

* Fraczkiewicz, R. and Braun, W. (1998) [J. Comp. Chem., 19, 319-333](https://doi.org/10.1002/(SICI)1096-987X(199802)19:3<319::AID-JCC6>3.0.CO;2-W).

### Details

Analysis of protein commonly requires the partition of their structure into regions such as the surface, interior, or interface. Whenever you are interested into analysing such a partition in a particular protein, the following function of the _ptm_ package may be of help:

* [acc.dssp](/acc.dssp)
* get.area (the current document)
* [atom.dpx](/atom.dpx)
* [res.dpx](/res.dpx)
* [stru.part](/stru.part)

The function **get.area()** returns the surface energies computed as described by  Fraczkiewicz & Braun. Further details can be obtained [here.](http://curie.utmb.edu/area_man.html) 

Using as example the human hexokinase 1, which is a monomeric protein, the output of the current function would look something like this:

```{r}
hHK1 <- get.area(pdb = '1cza')
kable(head(hHK1))
```
