---
title: "Scan a Protein in Search of Acetylation Sites"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Scans the indicated protein in search of acetylation sites

### Usage
ac.scan(up_id, db = 'all')

### Arguments
_up\_id_ a character string corresponding to the UniProt ID.
_db_ the database where to search. It should be one among 'PSP', 'dbPTM', 'all'.

### Value
Returns an dataframe where each row corresponds to an acetylable residue.

### References
Hornbeck et al. [Nucleic Acids Res. 2019 47:D433-D441](https://www.ncbi.nlm.nih.gov/pubmed/30445427).
Huang et al. [Nucleic Acids Res. 2019 47:D298-D308](https://www.ncbi.nlm.nih.gov/pubmed/30418626).

### See Also
_meto.scan()_, _p.scan()_, _me.scan()_, _ub.scan()_, _su.scan()_, _gl.scan()_, _sni.scan()_, _ni.scan()_, _ptm.scan()_, _reg.scan()_, _dis.scan()_

### Details

[Acetylation](https://www.sciencedirect.com/science/article/pii/S1570963916301169) is one of the major post-translational protein modifications in the cell, with manifold effects on the protein level as well as on the metabolome level. In recent years, our understanding of protein acetylation has increased significantly by global proteomics analyses and in depth functional studies. 

The package _ptm_ provides a function, **ac.scan()**, that aims to integrate information from different PTM databases, and assist to identify acetylation sites in a given protein. The UniProt ID of the protein of interest must be passed as argument. A second argument for this function is _db_, used to indicate the database to be searched. The databases accessible for this purpose are:

* [PSP](https://www.phosphosite.org): PhosphoSitePlus.
* [dbPTM](http://dbptm.mbc.nctu.edu.tw): database of Post-Translational Modifications.


By default, **ac.scan()** use all above databases to search for acetylation sites. In this case, it should be noted that the same acetylation site can appear twice if it is present in both databases. If we wish to get a list of unique acetylation sites found, for instance,  in the protein Histone H2B type 1-K, we can proceed as follows:

First, you need to search for the UniProt ID. If the protein in question is found into MetOSite, you can search the ID:

```{r}
id <- meto.list('Histone H2B type 1-K')$prot_id
id
```

Otherwise, go to the [UniProt](https://www.uniprot.org) page to search for the ID.

Then,

```{r}
unique(ac.scan(id)$modification)
```
