---
title: "Scan a Protein in Search of OGlcNAc Sites"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Scan the indicated protein in search of glycosylation sites

### Usage
gl.scan(up_id, db = 'all')

### Arguments
_up_id_	a character string corresponding to the UniProt ID.
_db_ the database where to search.

### Value
Returns a dataframe where each row corresponds to a modifiable residue.

### References
Hornbeck et al. [Nucleic Acids Res. 2019 47:D433-D441](https://www.ncbi.nlm.nih.gov/pubmed/30445427).
Huang et al. [Nucleic Acids Res. 2019 47:D298-D308](https://www.ncbi.nlm.nih.gov/pubmed/30418626).

### See Also
_ac.scan()_, _meto.scan()_, _p.scan()_, _me.scan()_, _ub.scan()_, _su.scan()_, _dis.scan()_, _sni.scan()_, _ni.scan()_, _ptm.scan()_, _reg.scan()_

### Details

[O-GlcNAcylation](https://www.frontiersin.org/articles/10.3389/fendo.2018.00599/full) is an important post-translational modification governed by a single pair of enzymes–O-GlcNAc transferase (OGT) and O-GlcNAcase (OGA). These two enzymes mediate the dynamic cycling of O-GlcNAcylation on a wide variety of cytosolic, nuclear and mitochondrial proteins in a nutrient- and stress-responsive fashion.

The package _ptm_ provides a function, **gl.scan()**, that assist in the identification of OGlcNAc sites in a given protein. The UniProt ID of the protein of interest must be passed as argument. For instance, to scan the protein vimentin, we can proceed as follows:

First, you need to search for the UniProt ID. If the protein in question is found into MetOSite, you can search the ID:

```{r}
id <- meto.list('Vimentin')$prot_id
id
```

Otherwise, go to the [UniProt](https://www.uniprot.org) page to search for the ID.

Then,

```{r}
gl.scan(id)
```

