---
title: "Scan a Protein in Search of S-Nitrosylation Sites"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Scans the indicated protein in search of nitration sites

### Usage
ni.scan(up_id)

### Arguments
_up_id_	a character string corresponding to the UniProt ID.

### Value
Returns a dataframe where each row corresponds to a modified residue.

### References
Huang et al. [Nucleic Acids Res. 2019 47:D298-D308](https://www.ncbi.nlm.nih.gov/pubmed/30418626).

### See Also
_ac.scan()_, _meto.scan()_, _p.scan()_, _gl.scan()_, _ub.scan()_, _su.scan()_, _dis.scan()_, _sni.scan()_, _me.scan()_, _ptm.scan()_, _reg.scan()_

### Details

[Protein nitration](https://www.ncbi.nlm.nih.gov/pubmed/23157446) represents a post-translational modification that occurs primarily on tyrosine residues. This modification disrupts nitric oxide signaling and skews metabolism towards pro-oxidant processes.

The package _ptm_ provides a function, **ni.scan()**, that assists in the identification of nitration sites in a given protein. The UniProt ID of the protein of interest must be passed as argument. For instance, to scan the mouse mitochondrial aspartate aminotransferase (P05202):

```{r}
ni.scan("P05202")
```

