---
title: "Scan a Protein in Search of S-Nitrosylation Sites"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Scans the indicated protein in search of S-nitrosylation sites

### Usage
sni.scan(up_id)

### Arguments
_up_id_	a character string corresponding to the UniProt ID.

### Value
Returns a dataframe where each row corresponds to a modifiable residue.

### References
Huang et al. [Nucleic Acids Res. 2019 47:D298-D308](https://www.ncbi.nlm.nih.gov/pubmed/30418626).

### See Also
_ac.scan()_, _meto.scan()_, _ni.scan()_, _gl.scan()_, _ub.scan()_, _su.scan()_, _dis.scan()_, _ptm.scan()_, _me.scan()_, _p.scan()_, _reg.scan()_

### Details

[S-Nitrosylation](https://www.ncbi.nlm.nih.gov/pubmed/15688001) is the covalent attachment of a nitric oxide group (-NO) to cysteine thiol within a protein to form an S-nitrosothiol (SNO). S-nitrosylation has diverse regulatory roles in bacteria, yeast and plants and in all mammalian cells.

The package _ptm_ provides a function, **Sni.scan()**, that assist in the identification of S-nitrosylation sites in a given protein. The UniProt ID of the protein of interest must be passed as argument. For instance, to scan the protein vimentin, we can proceed as follows:

First, you need to search for the UniProt ID. If the protein in question is found into MetOSite, you can search the ID:

```{r}
id <- meto.list('Vimentin')$prot_id
id
```

Otherwise, go to the [UniProt](https://www.uniprot.org) page to search for the ID.

Then,

```{r}
sni.scan(id)
```

