---
title: "Scan a Protein in Search of Sumoylation Sites"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Scans the indicated protein in search of sumoylation sites

### Usage
su.scan(up_id, db = 'all')

### Arguments
_up_id_	a character string corresponding to the UniProt ID.

_db_	the database where to search. It should be one among 'PSP', 'dbPTM', 'all'.

### Value
Returns a dataframe where each row corresponds to a modifiable residue.

### References
Hornbeck et al. [Nucleic Acids Res. 2019 47:D433-D441](https://www.ncbi.nlm.nih.gov/pubmed/30445427).
Huang et al. [Nucleic Acids Res. 2019 47:D298-D308](https://www.ncbi.nlm.nih.gov/pubmed/30418626).


### See Also
_ac.scan()_, _meto.scan()_, _ni.scan()_, _gl.scan()_, _ub.scan()_, _sni.scan()_, _dis.scan()_, _ptm.scan()_, _me.scan()_, _p.scan()_, _reg.scan()_

### Details

[Protein sumoylation](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3310159/) The post-translational modification by sumoylation is a major regulator of protein function that plays an important role in a wide range of cellular processes. 

The package _ptm_ provides a function, <font size="1">**su.scan()**</font>, that aims to integrate information from different PTM databases, and assist to identify sumoylation sites in a given protein. The UniProt ID of the protein of interest must be passed as argument. A second argument for this function is _db_, used to indicate the database to be searched. The databases accessible for this purpose are:

* [PSP](https://www.phosphosite.org): PhosphoSitePlus.
* [dbPTM](http://dbptm.mbc.nctu.edu.tw): database of Post-Translational Modifications.


By default, <font size="1">**su.scan()**</font> use all above databases to search for methylation sites. In this case, it should be noted that the same sumoylation site can appear twice if it is present in both databases. If we wish to get a list of unique ubiquitination sites found, for instance,  in the protein Histone H3.1t, we can proceed as follows:

First, you need to search for the UniProt ID. If the protein in question is found into MetOSite, you can search the ID:

```{r}
id <- meto.list('Histone H3.1t')$prot_id
id
```

Otherwise, go to the [UniProt](https://www.uniprot.org) page to search for the ID.

Then,

```{r}
unique(su.scan(id)$modification)
```

