---
title: "Scan a Protein in Search of Ubiquitination Sites"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Scans the indicated protein in search of ubiquitination sites

### Usage
ub.scan(up_id, db = 'all')

### Arguments
_up_id_	a character string corresponding to the UniProt ID.

_db_	the database where to search. It should be one among 'PSP', 'dbPTM', 'all'.

### Value
Returns a dataframe where each row corresponds to a modifiable residue.

### References
Hornbeck et al. [Nucleic Acids Res. 2019 47:D433-D441](https://www.ncbi.nlm.nih.gov/pubmed/30445427).
Huang et al. [Nucleic Acids Res. 2019 47:D298-D308](https://www.ncbi.nlm.nih.gov/pubmed/30418626).

### See Also
_ac.scan()_, _meto.scan()_, _ni.scan()_, _gl.scan()_, _su.scan()_, _sni.scan()_, _dis.scan()_, _ptm.scan()_, _me.scan()_, _p.scan()_, _reg.scan()_

### Details

[Protein ubiquitination](https://www.nature.com/articles/cr201639) is a dynamic multifaceted post-translational modification involved in nearly all aspects of eukaryotic biology. Ubiquitination can affect proteins in many ways: it can mark them for degradation via the proteasome, alter their cellular location, affect their activity, and promote or prevent protein interactions.

The package _ptm_ provides a function, **ub.scan()**, that aims to integrate information from different PTM databases, and assist to identify ubiquitination sites in a given protein. The UniProt ID of the protein of interest must be passed as argument. A second argument for this function is _db_, used to indicate the database to be searched. The databases accessible for this purpose are:

* [PSP](https://www.phosphosite.org): PhosphoSitePlus.
* [dbPTM](http://dbptm.mbc.nctu.edu.tw): database of Post-Translational Modifications.


By default, **ub.scan()** use all above databases to search for methylation sites. In this case, it should be noted that the same ubiquitination site can appear twice if it is present in both databases. If we wish to get a list of unique ubiquitination sites found, for instance,  in the protein Histone H3.1t, we can proceed as follows:

First, you need to search for the UniProt ID. If the protein in question is found into MetOSite, you can search the ID:

```{r}
id <- meto.list('Histone H3.1t')$prot_id
id
```

Otherwise, go to the [UniProt](https://www.uniprot.org) page to search for the ID.

Then,

```{r}
unique(ub.scan(id)$modification)
```

