---
title: "Building Up a Dataset for 3D Analyses"
author: "Juan Carlos Aledo"
date: "1/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

Suppose we are interested in addressing the questions: are the oxidable methionine sites found more often in the protein surface than in the protein core? How are MetO sites distributed among the different secundary structure elements? 

Addressing these questions involve the analysis of 3D data, which are often provided by the so-called PDB ( [Protein Data Bank](https://www.rcsb.org/)) files. Unfortunatly, in these PDB files we won't find information regarding those methionine sites liable to be oxidized. Such an information, can be found in other database named [*MetOSite*](https://metosite.uma.es). The aim of this vignette is to show how *ptm* can help us to combine the information provided by these two databases (PDB and MetOSite) to build up an adequate dataset to address issues such as those raised above.  

The functions from the package *ptm* that will be used herein are:

* meto.search()
* res.pdb()
* pep.pdb()

To focus our project a bit more, we will limit ourselves to study only those MetO sites found in human proteins after treatment with hydrogen peroxide. The relevant data can be obtained from the database [*MetOSite*](https://metosite.uma.es) using the *ptm* function `meto.search()`:

```{r}
data <- meto.search(organism = 'Homo sapiens', oxidant = 'H2O2')
```

This query returns over 4000 MetO sites. However, since all we want is to illustrate how to build a dataset containing structural (3D) information, an not so much to subsequently use such data (that is, we don't care about statistical power), we will reduce our set to a few (50 randomly choosen) MetO sites, which will shorten the analysis times.

```{r}
set.seed(123)
data <- data[sample(1:nrow(data), 50, replace = FALSE), 1:4]
head(data)
```








```{r}
res.pdb <- function(at, up, pdb, chain){
  a <- renum.pdb(pdb, chain, up)
  if(a$pdb[which(a$uni_pos == at)] == aa.at(at, up)){
    output <- TRUE
    attr(output, 'pdb_pos') <- a$pdb_pos[which(a$uni_pos == at)]
  } else {
    output <- FALSE
  }
  file.remove(paste(pdb, '.pdb', sep = ""))
  return(output)
}

pep.pdb <- function(pep, pdb){
  seq <- ptm::get.seq(pdb, db = 'pdb')
  if (gregexpr(toupper(data$env[i]), seq)[[1]] != -1){
      output <- TRUE
  } else {
      output <- FALSE
  }
  attr(output, 'at') <- gregexpr(toupper(data$env[i]), seq)[[1]]
  return(output)
}

# t <- pep.pdb('SHAGNKLAMQEF','2psn')
```


```{r message = FALSE, warning = FALSE}
data$pdb_pos <- data$pdb <- NA
for (j in 24:nrow(data)){
  print(j)
  up <- data$prot_id[j]
  pos <- data$met_pos[j]
  
  t <- uniprot2pdb(up) # list of PDB ID for this Uniprot entry
  if (nrow(t) > 0){
    for (i in 1:nrow(t)){
      pdb <- as.character(t$PDB[i])
      chain <- as.character(t$CHAIN[i])
    
      ## From metosite to uniprot
      a <- renum.meto(up)
      up_pos <- a$uni_pos[which(a$meto_pos == pos)]
    
      # ok <- res.pdb(up_pos, up, pdb, chain)
      ok <- pdb.res(up_pos, up, pdb, chain)
      if(ok){
        data$pdb[j] <- pdb
        # data$pdb_pos[j] <- attributes(res.pdb(up_pos, up, pdb, chain))$pdb_pos
        data$pdb_pos[j] <- attributes(pdb.res(up_pos, up, pdb, chain))$pdb_pos
        break
      } 
    }
  }
}
```

Now let's check that the MetO environment is well resolved in the PDB structure:

```{r message = FALSE, warning = FALSE}
data$env <- data$checked <- NA

for (i in 1:nrow(data)){
  print(i)
  if (!is.na(data$pdb[i])){
    ## From metosite to uniprot:
    a <- renum.meto(data$prot_id[i])
    up_pos <- a$uni_pos[which(a$meto_pos == data$met_pos[i])]
  
    ## Getting the MetO environment:    
    data$env[i] <- env.extract(prot = data$prot_id[i], 
                             db = 'uniprot',
                             c = up_pos,
                             r = 10)$Positive
    ## Getting the sequence from PDB
    seq <- ptm::get.seq(data$pdb[i], db = 'pdb')
    if (gregexpr(toupper(data$env[i]), seq)[[1]][1] != -1){
      data$checked[i] <- TRUE
    } else {
      data$checked[i] <- FALSE
    }
  }
}
```




## Bla bla bla
, for which a regulatory role has been described
For each site, the function returns information about: the Uniprot ID, the protein name, the position in the primary structure of the oxidized methionine, whether the oxidation has been reported to occur either _in vivo_, _in vitro_ or under both contitions..
However, since the two last features (species and oxidant) are the same for all the row in data, we can ignore these variables hereafter. On the other hand, since all we want is to illustrate how to build a dataset, an not so much to use such data, we will reduce 


