---
title: "aa.comp()"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
library(knitr)
```

### Description
Returns a table with the amino acid composition of the target protein

### Usage
aa.comp(target, uniprot = TRUE)

### Arguments

_target_	a character string specifying the UniProt ID of the protein of interest or, alternatively, the sequence of that protein.

_uniprot_	
logical, if TRUE the argument 'target' shoud be an ID.

### Value
Returns a dataframe with the absolute frequency of each type of residue found in the target peptide.

### See Also
_aa.at()_, _is.at()_, _renum.pdb()_, _renum.meto()_, _renum()_

### Details

The _ptm_ package offers a set of ancillary functions aimed to carry out rutinary work, which may be needed when more elaborated analysis are required. Among these ancillary function are:

* [aa.at](./aa.at)
* [is.at](./is.at)
* aa.comp (the current document)
* [renum.pdb](./renum.pdb)
* [renum.meto](./renum.meto)
* [renum](./renum)


The function **aa.comp()** provides us the amino acid composition of a given sequence.

```{r}
library(knitr)
mytable <- aa.comp('P01009')
mytable$percent <- round(100 * mytable$frequency/sum(mytable$frequency), 1)                
kable(mytable)
```

Above we have passed the UniProt ID of the protein of interest, but we can also provide any customized sequence:

```{r}
library(knitr)
mytable <- aa.comp('MAARESAPFDR', uniprot = FALSE)
mytable$percent <- round(100 * mytable$frequency/sum(mytable$frequency), 1)                
kable(mytable)
```



