---
title: "is.at()"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Checks if a given amino acid is at a given position

### Usage
is.at(at, target, aa = 'M', uniprot = TRUE)

### Arguments

_at_	the position in the primary structure of the protein.

_target_	a character string specifying the UniProt ID of the protein of interest or, alternatively, the sequence of that protein.

_aa_	the amino acid of interst.

_uniprot_	logical, if TRUE the argument 'target' shoud be an ID.

### Value
Returns a boolean. Either the residue is present at that position or not.

### See Also
_aa.at()_, _renum.pdb_, _renum.meto()_, _renum()_

### Details

The _ptm_ package offers a set of ancillary functions aimed to carry out rutinary work, which may be needed when more elaborated analysis are required. Among these ancillary function are:

* [aa.at](./aa.at)
* is.at (the current document)
* [aa.comp](./aa.comp)
* [renum.pdb](./renum.pdb)
* [renum.meto](./renum.meto)
* [renum](./renum)


Sometimes it may be convinient to check if the reside found at a given position is the one we expect, then we resort to the function **is.at()**. 

For instance when checking for data quality control in [MetOSite](https://metosite.uma.es/) (a database for methionine oxidation), we want to make sure that the annotation of the MetO site (at which position we find a sulfoxidized methionine) corresponds to a methionine residue in the original protein sequence. For example, if we query MetOSite for cytochrome c (P00004), we will find that Met80 is oxidized:

```{r}
meto.scan('P00004')$Metosites[[1]]
```

Now, let's chekc that certainly we find methionine at position 80 in the protein sequence:

```{r}
is.at(at = 80, target = get.seq('P00004', db = 'metosite'), uniprot = FALSE)
```
 
Note, that we have passed a string (the sequence of the mature cytochrome c) to the argument _target_:

```{r}
 get.seq('P00004', db = 'metosite')
```

Also, you may have noted that we didn't pass any value to the argument _aa_, that is because its value by default is 'M'.


