---
title: "List Proteins Found in MetOSite Matching a Keyword"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```


### Description
Lists proteins found in MetOSite with names matching the keyword.

### Usage
meto.list(keyword)

### Arguments
_keyword_ a character string corresponding to the keyword.

### Value
This function returns a dataframe with the UniProt id, the protein name and the species, for those proteins present into MetOSite whose name contains the keyword.

### References
Valverde et al. 2019. MetOSite: an integrated resource for the study of methionine residues sulfoxidation, [Bioinformatics 35:4849-4850](https://doi.org/10.1093/bioinformatics/btz462).

### See Also
_meto.search()_, _meto.scan()_

### Details

We have search the literature to find those protein-bound methionine residues that have been detected, either _in vitro_ or _in vivo_, as MetO. All this information has been collected and published in the form of a database: [MetOSite](https://metosite.uma.es).

Thus, MetOSite is a database that provides mannually curated data related to experimentally confirmed sulfoxidation sites and [the effect of such PTM on the protein properties](https://www.ncbi.nlm.nih.gov/pubmed/31359525). Currently, it contains over 7300 sites found in more than 3500 different proteins from over 20 species. 

The _ptm_ packages offers three functions that help to explore and download data from MetOSite. 

* meto.list (the current tutorial)
* [meto.scan](./meto.scan)
* [meto.search](./meto.search)


Thus, if you wish to know whether your protein of interest has been described to contain MetO sites, the function **meto.scan()** will help you.

If, on the contrary, you're interested on sets of proteins or MetO sites, grouped according to criteria such as functional effect, species or oxidants, then the place to go is **meto.search()**

Finally, the function **meto.list()** provides a list with the proteins found in MetOSite whose names contain the keyword provided as argument to the function. 

For instance, using the word _calcium_ as keyword:

```{r}
ca <- meto.list('calcium')
head(ca)
```

In this way, the object we have named *ca* is a dataframe that contains tens of proteins (one per row) whose names include the word *calcium*. Now, if you wish, we can further explore any of them using **meto.scan()**. For instance, let's say we are interested in the protein "Calcium/calmodulin-dependent protein kinase type II subunit delta", which is found in the row 11 from the dataframe *ca* :

```{r}
meto.scan(ca$prot_id[11])$Note
```
Note that we have only requested the 'Note' field. For a more detailed report, please check the documentation of the function **meto.scan()**.

