---
title: "gogo()"
author: "Juan Carlos Aledo"
date: "24/10/2020"
output: html_document
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
library(knitr)
```


Let's focus on the function hdfisher.go(). Since this function carries out a hypothesis-driven GO term enrichment analysis, the first thing we need to start with is a working hypothesis. Let's suppose that we hypothesize that among all the protein having at least one methionine residue whose oxidation affect the protein properties (this is our background set), those that are activated as consequence of methionine sulfoxidation (this is our target set) are enriched in proteins related to calcium metabolism (this is our hypothesis).  

First, we start getting the dataframe that we will use as background. For this end, we will use the function background.go(): 

```{r}
sites <- meto.search(highthroughput.group = FALSE,
                 bodyguard.group = FALSE,
                 regulatory.group = TRUE)
bs <- background.go(unique(sites$prot_id))
```

Then we form the target set:

```{r}
gsites <- meto.search(highthroughput.group = FALSE,
                 bodyguard.group = FALSE,
                 regulatory.group = TRUE,
                 gain.activity = 1)
target <- unique(gsites$prot_id)
```

Finally, we are ready to use the function hdfisher.go():

```{r}
enrichment <- hdfisher.go(target = target, 
                           background = bs, 
                           query = "calcium")
```

Now we can request the contingency table:

```{r}
enrichment[[1]]
```

and the p-value:

```{r}
enrichment[[2]]
```

Note that the hdfisher.go() function compute the p-value for an enrichment test by default. However, we can also address depletion hypothesis. For instance, if we hypothesize that protein related to the RNA metabolism are underrepresented in the background set, se can set the "analysis" argument to "depletion":

```{r}
depletion   <- hdfisher.go(target = target, 
                           background = bs, 
                           query = "RNA", 
                           analysis = "depletion")
depletion[[1]]
depletion[[2]]
```

