---
title: "id.features()"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Obtains features related to the provided id

### Usage
id.features(id, features = "")

### Arguments
_id_	the UniProt identifier of the protein of interest.

_features_	a string identifying the features (comma separated) to be recovered.

### Value 
Returns a named list with the requested features. By default the the function provides info regarding the following features: id, reviewed, entry name and organism. If wished, this list of features can be expanded using the argument 'features'.

### Details

The _ptm_ package offers a set of functions aimed to assist us to download and handle sequences from different databases: 

* [get.seq] (./get.seq)
* [prot2codon](./prot2codon)
* [id.mapping](./id.mapping)
* id.features (current tutorial)
* [species.mapping](./species.mapping)

Herein we illustrate the use of **id.features()** using the UniProt ID P35557, which corresponds to the human glucokinase (hexokinase-4)

```{r}
gk <- id.features('P35557')
gk
```
By default we only recover information regarding a few basic features, but we can expand the information retrieved by this function using the argument _features_. For instance,

```{r}
gk <- id.features('P35557', features = 'ec,keywords,database(PDB)')
```

Note, that we have requested three additional features:

The Enzyme Commission number for this enzyme
```{r}
gk$EC
```

Keywords linked to this protein
```{r}
gk$Keywords
```

```{r}
gk$`Cross-reference_(PDB)`
```
There is a larga list of features that can be retrieved. You can look up your relevant feature's name in the full list of UniProtKB found [herein](https://www.uniprot.org/help/uniprotkb_column_names).

Finally, one last example. Let's find the subcellular locations for this protein:

```{r}
gk <- id.features('P35557', features = 'comment(SUBCELLULAR LOCATION)')
gk$`Subcellular_location_[CC]`
```

