---
title: "pairwise.dist()"
# author: "Juan Carlos Aledo"
# date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ptm)
```

### Description
Computes the pairwise distance matrix between two sets of points

### Usage
pairwise.dist(a, b, squared = TRUE, ...)

### Arguments
_a, b_	matrices (NxD) and (MxD), respectively, where each row represents a D-dimensional point.

_squared_	return containing squared Euclidean distance.

### Value
Euclidean distance matrix (NxM). An attribute "squared" set to the value of param squared is provided.

### See Also
_res.dist()_, _dist2closest()_, _ball()_

### Details

The continuous formation and rupture of non-covalent bonds (intra- and intermolecular), gives rise to dynamic molecular interactions, which are cosubstantial to life. These non-covalent interactions can adopt many forms, but all of them are fundamentally electrostatic in nature, and strongly influenced by distances.

| Type of interaction   | Energy dependence on the distance |
|-----------------------|-----------------------------------|
| Ion-ion               | $1/r$                             |
| Ion-dipole            | $1/r^2$                           |
| Dipole-dipole         | $1/r^3$                           |
| Ion-induced dipole    | $1/r^4$                           |
| Dipole-indiced dipole | $1/r^5$                           |
| Dispersion            | $1/r^6$                           |


Therefore, computing spatial distances (in ångströms, Å), either between atoms or residues, is a useful task in many different contexts. A number of function from the package _ptm_ will help us in these tasks:

* [res.dist](./res.dist)
* [dist2closest](./dist2closest)
* [ball](./ball)
* pairwise.dist (the current document)


The most elementary of all these functions is, **pairwise.dist()**, which takes two sets of points (in the form of matrices) tailored by the user, and returns a distance matrix. Let's just   

```{r}
a <- matrix(c(1,1,1,1,0,0,1,0,1), ncol = 3, byrow = TRUE)
rownames(a) <- c('point-1', 'point-2', 'point-3')
b <- matrix(c(0,0,0,1,1,1), ncol = 3, byrow = TRUE)
rownames(b) <- c('point-4', 'point-5')
a
b
```

and now let's compute the distances between the two sets of points:

```{r}
pairwise.dist(a, b, square = FALSE)
```
